Master's Project Journal

### 2019-04-18 thesis/NetworkImplemenation.tex
per
62

Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope. Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure. Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$. One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected. $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
%TODO diagram or picture of some sort

\section{\acs{ADPLL} Characterisation}
%TODO why each PLL

@@

\section{Minor Variations}\label{section:minor_variations}
%TODO osc size

\section{\acs{ADPLL} Network Implementation}
\subsection{2x2 Network}
\subsection{3x3 Network}

>>>>

Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope. Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure. Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$. One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected. $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
%TODO diagram or picture of some sort

\section{\acs{ADPLL} Characterisation}
Before combining the \ac{ADPLL}s in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance. These tests were carried out with the \ac{ADPLL}s locked to an external reference at $5~\si{\mega\hertz}$. As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown. Jitter was calculated based on the performance on nine individual \ac{ADPLL}s all locked simultaneously to the $5~\si{\mega\hertz}$ external reference. While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.

The \label{table:characterisation} contains the results of this analysis. The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \ac{ADPLL}s implementing \ac{RO}s are identical. As the lock range would be expected to be slightly wider than the capture range in a design limited by the loop filter, it can be deduced that the limiting factor in the tuning range of the \ac{ADPLL} is the \ac{DCO} frequency range. This is confirmed by a cursory comparison against the values in Table \ref{table:adpll2}, where the \ac{RO} range is given as $4.571-5.518~\si{\mega\hertz}$. This range of $0.947~si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$ which confirms the \ac{RO} is the limiting factor. In \ac{ADPLL} 1 however, the lock and capture ranges show some difference, with the lock range being slightly wider. The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, and $1.08~\si{\mega\hertz}$ is the point that the lock was observed to be lost.

\begin{table}[]
    \begin{centre}
        \begin{tabular}{lrrr}
        \multicolumn{1}{c}{}          & \multicolumn{1}{c}{ADPLL Design 1}  & \multicolumn{1}{c}{ADPLL Design 2} & \multicolumn{1}{c}{ADPLL Design 3} \\ \cline{2-4}       
        Capture Range                 & $1.08-8.61~\si{\mega\hertz}$       & $4.869-5.877~\si{\mega\hertz}$    & $4.377-5.363~\si{\mega\hertz}$       \\
        Lock Range                    & $1.08-8.635~\si{\mega\hertz}$      & As above                           & As above                            \\
        Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
        Skew (Mean \acs{TIE} to Ref.) & $7.7939~\si{\nano\second}$         & $6.2101~\si{\nano\second}$         & $5.8145~\si{\nano\second}$          \\
        Maximum \acs{TIE}             & $8.7847~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
        $k_p$                         & $\frac{1}{32}$                     & $\frac{1}{16}$                     & $\frac{1}{16}$                      \\
        $k_i$                         & $\frac{1}{256}$                    & $\frac{1}{64}$                     & $\frac{1}{64}$  
        \end{tabular}
    \end{centre}
    \caption[\ac{ADPLL} Design Comparison]{\ac{ADPLL} Design Comparison.}
    \label{table:characterisation}
\end{table}

Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step is has the worst jitter, by a significant amount. The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected.

@@

\section{Minor Variations}\label{section:minor_variations}

\subsection{Impact of Gain Variation}
The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error. To large a proportional gain will result in overly aggressive overcompensation, thus negatively affecting the jitter. Eventually the violence of the jitter will result in a loss of lock. However, if the gain is too small, the frequency or period adjustment will not be sufficiently large to cause the \ac{ADPLL} to lock. Similarly, appropriate values of $k_i$ are required, and the relationship between proportional and integral gains must satisfy the conditions set out by Koskin \cite{koskin2018generation}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../fixed_kp.png}
    \caption[Fixed $k_p$ integral gain sweeps]{\Fixed $k_p$ integral gain sweeps.}
    \label{fig:gain_sweep}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../fixed_kp_zoom.png} 
    \caption[Fixed $k_p$ integral gain sweeps]{\Fixed $k_p$ integral gain sweeps.}
    \label{fig:gain_sweep_zoom}
\end{figure}

Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains. It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$. These distinct sections of the plot represent locked and unlocked operation respectively. Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$.
\begin{table}[!h]
    \begin{center}
        \begin{tabular}{lrr}
            \multicolumn{1}{c}{$k_p$} & \multicolumn{1}{c}{$k_i$} & \multicolumn{1}{c}{Ratio} \T\B\\
            \hlines
            $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$                     \T\B\\
            $\frac{2}{16}$            & $\frac{10}{64}$           & $0.8$                     \T\B\\
            $\frac{3}{16}$            & $\frac{12}{64}$           & $1$                       \T\B\\
            $\frac{4}{16}$            & $\frac{14}{64}$           & $1.1428$                  \T\B
        \end{tabular}
    \end{center}
    \vspace{-0.5cm}
    \caption[Loss of Lock Gains]{Loss of Lock Gains.}
    \label{table:lolgains}
\end{table}

Despite the relatively small number of proportional gains, it is clear that the relationship set out by Koskin does not hold in this system, but rather a scaled version of it.  This difference is attributable to a difference in \ac{LF} structure. In his simulations the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests. The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter. The removal of this register causes a change in the relationship between proportional and integral gains, a change that will be investigated further in a later section.

\subsection{Distribution of Period Steps}
To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined. The best way to visualise this distribution is through a histogram, placing periods of in a certain range into the same bin. This can be visualised effectively using Matlab's \texttt{histplot}. In order to obtain data, three \ac{ADPLL}s implementing both the \ac{FPGA} clocked and inverter based \ac{DCO}s were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
\begin{figure}[h]%
    \centering
    \subfloat[Inverted delay based \ac{RO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_ring} }}%
    \subfloat[\ac{FPGA} clocked \ac{DCO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_pa} }}%
    \caption[Example distribution of periods]{Example distribution of periods.}    
    \label{fig:dists}
\end{figure}
Figure \ref{fig:dists} contains the output of \texttt[histplot} for each of the three oscillators superimposed in one plot for each \ac{DCO} type. As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other thereby making the plot difficult to read. In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters to remove was altered from twice the control code to 16 times the control code. This results in a more pronounced separation of period steps while being indicative of the variation due to implementation.
The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network. If this analysis was carried out for the period steps, similar behaviour would be observed.

%TODO removal of delay
%TODO osc size

\section{\acs{ADPLL} Network Implementation}
%TODO why use each PLL
\subsection{2x2 Network}
\subsection{3x3 Network}
